---
title: "An Introduction to Transparency: West Bountiful City"
author: "Michael Jensen"
output: html_notebook
---
# Connect to AWS

```{r}
library(odbc)

# "transpAWS" is what I've named my connection to AWS. Yours may be different.
odbc_aws <- dbConnect(odbc::odbc(), "transpAWS")
```

# What tables exist in the Transparency database?

```{sql, connection=odbc_aws, output.var=table_names_of_tables}
SELECT TABLE_NAME
FROM information_schema.tables
WHERE TABLE_SCHEMA = "transparency"
```

# What is West Bountiful City's Transparency ID?

```{r}
# Query names and IDs from the entity table:
table_entity <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT name, id
    FROM entity")
```

West Bountiful's Transparency ID: 623

# Who uploads data for West Bountiful?

```{r}
usernames <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT DISTINCT upload_username, upload_date
    FROM batch
    WHERE entity_id = 623
    ORDER BY upload_date DESC")
```

# What's an overview of the files West Bountiful has uploaded?

```{r}
table_batch <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT 
      id, 
      upload_date, 
      processed_date,
      file_record_count,
      file_total_amount,
      begin_txn_date,
      end_txn_date,
      status,
      status_message
    FROM batch
    WHERE entity_id = 623
    ORDER BY upload_date DESC")
```

# What expense transactions are contained in the most recently processed batch, ordered from highest to lowest?

```{r}
batch_id <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT id
    FROM batch
    WHERE entity_id = 623
    AND status = 'PROCESSED'
    ORDER BY processed_date DESC
    LIMIT 1")

batch_id <- batch_id$id[[1]]

table_transactions <- 
  dbGetQuery(
    odbc_aws,
    paste("
          SELECT 
            posting_date,
            amount,
            fiscal_year,
            entity_trans_id,
            description,
            org1, org2, org3,
            cat1, cat2, cat3,
            fund1, fund2,
            vendor_id
          FROM transaction
          WHERE batch_id = ", batch_id, "
          AND type = 1
          ORDER BY amount DESC"))  # type = 1: expense, type = 2: revenue
```

# What does org1, org2, org3, cat1, etc. mean?

```{r}
table_transaction_group <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT id, name
    FROM transaction_group")
```

```{r}
table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("org1" = "id")) %>% 
  select(-org1) %>% 
  rename(org1 = name)

table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("org2" = "id")) %>% 
  select(-org2) %>% 
  rename(org2 = name)


table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("org3" = "id")) %>% 
  select(-org3) %>% 
  rename(org3 = name)

table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("cat1" = "id")) %>% 
  select(-cat1) %>% 
  rename(cat1 = name)


table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("cat2" = "id")) %>% 
  select(-cat2) %>% 
  rename(cat2 = name)


table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("cat3" = "id")) %>% 
  select(-cat3) %>% 
  rename(cat3 = name)


table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("fund1" = "id")) %>% 
  select(-fund1) %>% 
  rename(fund1 = name)


table_transactions <- 
  table_transactions %>% 
  left_join(table_transaction_group,
            by = c("fund2" = "id")) %>% 
  select(-fund2) %>% 
  rename(fund2 = name)
```

# Who are the vendors for each transaction?

```{r}
table_vendor <- 
  dbGetQuery(
    odbc_aws,
    "
    SELECT id, name
    FROM vendor
    WHERE entity_id = 623")

table_transactions <- 
  table_transactions %>% 
  left_join(table_vendor,
            by = c("vendor_id" = "id")) %>% 
  select(-vendor_id) %>% 
  rename(vendor_id = name)
```

# Close

```{r}
dbDisconnect(odbc_aws)
```